---
title: "Untitled"
author: "Dániel Vörös"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=FALSE, message=F}
# set this option in the first code chunk in the document
knitr::opts_chunk$set(echo = F, cache = T)

# libraries
library(geodata)
library(tidyr)
library(ggplot2)
```

# Climate

## Getting climate data

-   `geodata` package for downloading CMIP6 climate data
-   Tons of models (e.g. "CNRM-CM6-1-HR")
-   Socio - economic Pathways: 126, 245, 370 or [585]{.underline}
-   predictions for different time-intervals: 2021-2040, 2041-2060 or [2061-2080]{.underline}
-   monthly resolution: minimal-, maximal temperature, precipitation

```{r}
#f <- "*EUR-44*"
#UDG.datasets(pattern = f)

#bio10 <- cmip6_world("CNRM-CM6-1-HR", "585", "2061-2080", var="tmin", res=10, path=tempdir())
biomin <- cmip6_tile(15, 49, "CNRM-CM6-1-HR", "585", "2061-2080", var="tmin", res=10, path=tempdir())
#bioavg <- cmip6_tile(15, 49, "CNRM-CM6-1-HR", "585", "2061-2080", var="tavg", res=10, path=tempdir())
biomax <- cmip6_tile(15, 49, "CNRM-CM6-1-HR", "585", "2061-2080", var="tmax", res=10, path=tempdir())

#b <- bio10$tmin01
#bc <- crop(b, ext(-30, 50, 30, 80))
plot(crop(biomax, ext(-11, 35, 42, 71)))
```

## In time

```{r}
#plot(biomax)

#which(!is.nan(values(bio$wc2.1_30s_tmin_01)))
#barplot(unlist(biomin[1447]))
#barplot(unlist(biomax[1447]))


b <- cbind( (biomin[1447]), (biomax[1447]) )
b2 = b |> 
  #select( !starts_with("no_A", ignore.case = F)) |> 
  pivot_longer(cols= everything(), names_to=c("parameter", "month"), values_to = "val", names_pattern = "wc2.1_30s_(.*)_(.*)")

b3 <- b2 |> pivot_wider(names_from = parameter, values_from = val)

ggplot(b3, aes(x=month))+
  geom_linerange(aes(ymin=tmin, ymax=tmax))

```

## How to model temperature?

-   Sinus function in literature

```{r}
func <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dX <-  ( (20*cos(t)+10)  -X)
    w = t > et$min & t < et$max
    if( any(w) ) {
      #cat(paste(round(t, 3), round(X, 2), round(dX, 2), "\n") )
      dX <- dX * et[w, "times"]
    }
    list(dX)
  })
}
```


```{r}
state <- c(X = 10)
times <- seq(0, 20, by = 0.01)
no_errors = 0
length=0.1

<< eval_ode >>

```


```{r eval_ode, eval=F}
at = runif(no_errors, 0, max(times))
a = rnorm(length(at), sd=6)
parameters <- list(et = data.frame(min=at-length/2, max=at+length/2, times=a))

out <- ode(y = state, times = times, func = func, parms = parameters)

ggplot(as.data.frame(out), aes(x=time, y=X))+
  geom_line()+
  labs(x="Time", y="Temperature")

```


## Temperature deviations

```{r}
set.seed(1)

state <- c(X = 10)
times <- seq(0, 20, by = 0.01)
no_errors = 4
length=0.1

<< eval_ode >>

```


# Model

## Equations

$$\frac{dN_g}{dt1}=N_g r(T) \left( 1-\frac{\sum_{i=1}^{G}N_i}{K} \right) - N_g p_{sleep} + D_g p_{wake}\\
\frac{dD_g}{dt}= N_g p_{sleep} - D_g p_{wake} - D_g \delta \\
r(T)= \dots $$

<!--
$$x(T)=P_{opt} \frac{(T-T_{min})(T-T_{max})}{(T-T_{min})(T-T_{max})-(T-T_{opt})^2}$$
-->

```{r, eval=F}
Popt=1
Tmin=10
Trange=30
Topt=10
Tmax=Tmin+Trange

curve(Popt * ( (T-Tmin)*(T-Tmax) ) * ( (T-Tmin)*(T-Tmax)-(T-Topt)^2 ), xname="T", from=Tmin, to=Tmax)
abline(h=0)
gorbe <- function(T) return(Popt * ( (T-Tmin)*(T-Tmax) ) * ( (T-Tmin)*(T-Tmax)-(T-Topt)^2 ))
```

<!--
$$x(T, M)=e^{I} M^S e^{E_a \frac{T_0-T+T_{0K}}{k_B (T+T_{0K}) T_0 }}$$ $$x(T, M)=M^S e^{I + E_a \frac{T_0-T+T_{0K}}{k_B (T+T_{0K}) T_0 }}$$ 
-->

```{r, eval=F}
Sx    = -0.31;          # rate specific scaling coefficient
Ix    = -16.54;         # rate specific constant 
Eax   = -0.69;          # activation energy (eV)
T0K   = 273.15;         # 0 degres in Kalvin
Boltz = 8.617*10^(-5);  # Boltzmann constant (eV/K)
T0    = 293.15;         # normalization temperature (K)


x2<-function(Temp,My){
  x2= exp(Ix)*My^Sx*exp(Eax*(T0-(Temp+T0K))/(Boltz*(Temp+T0K)*T0));
}


curve(x2(x, 1), from=10, to=30)
```

```{r, eval=F}
curve(gorbe(x) - x2(x, 1) , from=10, to=30)
curve(x2(x, 1), add=T)

```

```{r, eval=F}
Trange=50
Tmin=10
a=1
b=0.25
Tmax=Tmin+Trange

curve(a*exp(b*T)*(Tmax-T)*(T-Tmin), from=Tmin, to=Tmax, xname="T", sub=paste("a", a, "b", b), lwd=2)
```


```{r, eval=F}
#b= 10/Trange

Trange=50*2
Tmin=10
a=1
b=10/Trange
#b=1.9
Tmax=Tmin+Trange

par(oma=c(0,0,0,3))
xxx <- curve(a*exp(b*T)*(Tmax-T)*(T-Tmin), from=Tmin, to=Tmax, xname="T", sub=paste("a", a, "b", b), lwd=2)
par(new=T)
curve(a*exp(b*T), add=T, xname="T", col="red", axes=F)
axis(4, col="red")
par(new=T)
curve((Tmax-T)*(T-Tmin), xname="T", add=T, col="blue", axes=F)
axis(4, col="blue", outer=T)

sprintf("%e", max(xxx$y))
```

```{r, eval=F}
Trange=11
Tmin=10
a=1
#b=100/Trange
b=1.9
Tmax=Tmin+Trange

par(oma=c(0,0,0,3))
xxx <- curve(a*exp(b*(T-Tmin)/Trange)*(Tmax-T)*(T-Tmin), from=Tmin, to=Tmax, xname="T", sub=paste("a", a, "b", b), lwd=2)
par(new=T)
curve(a*exp(b*T), add=T, xname="T", col="red", axes=F)
axis(4, col="red")
par(new=T)
curve((Tmax-T)*(T-Tmin), xname="T", add=T, col="blue", axes=F)
axis(4, col="blue", outer=T)

sprintf("%e", max(xxx$y))
```

```{r, eval=F}
plot(c(1,25,50,100, 200), c(8e-1,5e2, 2e3, 8e3, 3e4), xlab="Trange", ylab="max")
```

<!--
$$-(T_{range} e^{b \frac{T-T_{min}}{T_{range}}} (b^2 (T_{min} - T) (T_{min} + T_{range} - T) + b T_{range} (2 T_{min} + T_{range} - 2 T) + 2 T_{range}^2))/b^3 + c$$
-->


## Replication rate

$$c(b, T_{range}) = \int_{T_{min}}^{T_{max}} e^{b \frac{T-T_{min}}{T_{range}}} (T_{max}-T)(T-T_{min}) dT=  \frac{2 + b + (b - 2) e^b}{b^3} {T_{range}}^3$$ $$r(T) = A e^{b \frac{T-T_{min}}{T_{range}}} (T_{max}-T)(T-T_{min}) / c$$

```{r}
ff <- function(T, Tmin, Tmax, Trange, b, compensation){
  exp(b * (T-Tmin)/Trange) * (Tmax - T) * (T - Tmin) / compensation
}
ff2 <- function(T, Tmin, Tmax, Trange, b, compensation){
  exp(b/Trange)^(T-Tmin) * (Tmax - T) * (T - Tmin) / compensation
}
```

```{r plot_T, eval=F}
Tmax=Tmin+Trange
compensation = (2+b+(b-2)*exp(b))*Trange^3/b^3/A

ggplot()+
  stat_function(fun = ff, args=list(Tmin=Tmin, Tmax=Tmax, Trange=Trange, b=b, compensation=compensation), xlim = c(Tmin,Tmax), geom="area", fill="lightblue")+
  geom_function(fun = ff, args=list(Tmin=Tmin, Tmax=Tmax, Trange=Trange, b=b, compensation=compensation), xlim = c(Tmin-Trange/10,Tmax+Trange/10), color="blue")+
  labs(x="Temperature", y = "Replication rate")

#curve(ff(T, Tmin, Tmax, Trange, b, compensation), xname="T", from=Tmin-Trange/10, to=Tmax+Trange/10)
#abline(h=0)
```

```{r}
Trange=11
Tmin=10*2
A=1
b=1.9

<< plot_T >>
```














```{r, eval=F}
ff3 <- function(diff1, diff2, base, compensation){
  base^diff1 * diff2 * diff1 / compensation
}

Trange=11
Tmin=10*2
A=1
b=1.9

Tmax=Tmin+Trange
compensation = (2+b+(b-2)*exp(b))*Trange^3/b^3/A
base <- exp(b/Trange)

curve(ff3(diff1 = T-Tmin, diff2=Tmax-T, base, compensation), xname="T", from=Tmin-Trange/10, to=Tmax+Trange/10)
abline(h=0)

#spared: -, /
```

```{r}
#A=-b
#B=b*(Tmax+Tmin)/2-2
#C=Tmax+Tmin-b*Tmax*Tmin
#o = (-B-sqrt(B^2-4*A*C))/(2*A)
#abline(v=o)
```



